\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{amsmath,amssymb}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{enumitem}

\hypersetup{hidelinks}
\captionsetup{font=small,labelfont=bf}
\graphicspath{{figures/}}

\title{Predicting High Income from Census Attributes}
\author{Adult Income ML Study}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage

\textbf{A supervised machine-learning analysis of the UCI Adult / Census Income dataset}

\begin{quotation}
\textbf{Viewing figures:} Image paths are relative to this file (\texttt{./figures/...}). Open the preview from \texttt{reports/report.md} (not the repo root). If figures are missing, run \texttt{python scripts/03\_run\_eda.py} and later pipeline scripts, or \texttt{make eda} through \texttt{make report}.
\end{quotation}


\section{Introduction}

Income classification from census and survey variables is a longstanding benchmark in applied machine learning. The UCI Adult (Census Income) dataset records demographic, education, employment, and financial attributes for tens of thousands of individuals, with a binary label indicating whether annual income exceeds \$50,000. Predicting this label is useful as a methodological test bed for tabular classification, class imbalance, model comparison, and fairness auditing—not as a tool for individual-level decisions without further validation.

This report presents an end-to-end study: reproducible data preparation, exploratory analysis, comparison of classical and neural models, held-out evaluation, interpretability analysis, subgroup fairness diagnostics, calibration assessment, and an extension that removes sensitive attributes. Evidence is organised around a fixed traceability matrix linking each claim to tables, figures, or pipeline artifacts (see Appendix).


\section{Research Focus}

\subsection{Main research question}

\textbf{To what extent can income above \$50K/year be predicted from demographic, education, employment, and financial attributes in the Adult / Census Income dataset, and how do model family, preprocessing choices, calibration, interpretability, and subgroup fairness diagnostics affect the strength and reliability of that prediction?}

\subsection{Supporting objectives}

\begin{enumerate}
  \item Build a leakage-aware, reproducible binary classification dataset (DATA-001, DATA-002).
  \item Compare classical models and neural MLP variants under the same split and metric protocol (MODEL-001–005).
  \item Evaluate whether stronger predictive performance trades off against interpretability or subgroup disparity (INT-\textit{, FAIR-}).
  \item Produce an evidence-first report with traceable figures and tables (RPT-001).
\end{enumerate}

\subsection{Model selection criterion}

Models are compared using \textbf{mean 5-fold cross-validated macro F1 on the training split only}. The held-out test set is used once for final evaluation of the selected model (EVAL-001). Hyperparameters are tuned via randomized search on the training data; the test set is never used for selection or tuning.

\end{document}
